Examining spatial equity and accessibility to a public bicycle share program using a balanced floating catchment area approach

### 2021-03-19 Bike-share-spatial-equity/Bike-share-spatial-equity.Rmd
Revised draft with complete calculations

```{r sobi-stations-in-hamilton, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "The spatial distribution of bike share stations in Hamilton, Ontario. The service area of Hamilton Bike Share is outlined in blue and the city's downtown core is outlined in light blue. Dissemination areas within the Hamilton CMA that touch the bounding box of the core service area are outlined in grey."}
ggplot() + 
  geom_sf(data = hamilton_da_2016 %>% 
            dplyr::filter(sobi_service == "Yes"),
          color = "gray") +
  geom_sf(data = sobi_service, 
          fill = NA,
          color = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = "dark green",
          color = "dark green") +
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional" & 
                            hub_status == "Active"), # Remove deactivated hubs
          size = 1.5, 
          shape = 21, 
          fill = "blue") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

@@

## Accessibility Calculations at Different Thresholds

Consensus regarding the distance that individuals are willing to walk to access a PBSP is lacking, therefore the literature on walking behaviour was consulted to determine the thresholds for the sensitivity analysis. Previous studies have found that living within 250 metres [@fullerUseNewPublic2011] and 300 metres [@kabraBikeShareSystemsAccessibility2020] is correlated with bike share use, while other research has found that walking trips are less than 600 metres and rarely more than 1200 metres [@millwardActivetransportWalkingBehavior2013] or a median distance of 650 metres [@larsenQuarterMileReexamining2010]. For this reason, we experiment with various walking thresholds by conducting a sensitivity analysis to calculate accessibility at different walking times: 3 minutes, 5 minutes, 10 minutes, and 15 minutes. We repeat these calculations for both the current system and the original system to compare how accessibility changes at different thresholds with the addition of the equity stations.

```{r join-population-to-travel-time-data, echo=FALSE}
# current system with equity stations
ttm_walk <- left_join(ttm_walk, 
                      population_50x50 %>%
                        st_drop_geometry(),
                      by = "UID")
```

@@

```{r echo=FALSE}
#data frames for current and original systems
names(ttm_walk) <- c("UID", 
                     "hub", 
                     "travel_time",
                     "hub_type",
                     "hub_status",
                     "population",
                     "racks")
```

### Varying thresholds

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 3 minutes:
```{r}
results_3min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

All stations:
```{r}
results_3min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

Sanity check:
```{r}
sum(results_3min_conventional$los$los)
sum(results_3min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_3min_all$los$los)
sum(results_3min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 5 minutes:
```{r}
results_5min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

All stations:
```{r}
results_5min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

Sanity check:
```{r}
sum(results_5min_conventional$los$los)
sum(results_5min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_5min_all$los$los)
sum(results_5min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 10 minutes:
```{r}
results_10min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

All stations:
```{r}
results_10min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

Sanity check:
```{r}
sum(results_10min_conventional$los$los)
sum(results_10min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_10min_all$los$los)
sum(results_10min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 15 minutes:
```{r}
results_15min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

All stations:
```{r}
results_15min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

Sanity check:
```{r}
sum(results_15min_conventional$los$los)
sum(results_15min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_15min_all$los$los)
sum(results_15min_all$accessibility$accessibility)
```

Extract results:
```{r, include=FALSE}
los <- rbind(data.frame(results_3min_all$los, system = "All stations", threshold = "3 min"),
             data.frame(results_3min_conventional$los, system = "Conventional", threshold = "3 min"),
             data.frame(results_5min_all$los, system = "All stations", threshold = "5 min"),
             data.frame(results_5min_conventional$los, system = "Conventional", threshold = "5 min"),
             data.frame(results_10min_all$los, system = "All stations", threshold = "10 min"),
             data.frame(results_10min_conventional$los, system = "Conventional", threshold = "10 min"),
             data.frame(results_15min_all$los, system = "All stations", threshold = "15 min"),
             data.frame(results_15min_conventional$los, system = "Conventional", threshold = "15 min"))

@@

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 3 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "3 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 5 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "5 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 10 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "10 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 15 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "15 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

## Comparison

Include table with total # of people who can reach a station within 3 minutes walk, 5 minutes, 10 minutes, and 15 minutes, to compare accessibility at different thresholds for both current and original system.

Discussion
===================

Our analysis demonstrates that the stations added to Hamilton's public bicycle share program to increase equity for disadvantaged neighbourhoods achieved their goal by increasing accessibility, albeit only modestly.

By implementing equity stations in more areas with lower median household income in Hamilton, the PBSP has achieved greater horizontal equity by extending the spatial distribution of bicycles across the city. Under the the conditions of the original system, which did not have equity stations, the level of service was _X_. With the addition of twelve equity stations, the level of service increased to _X_.

However, we found that dissemination areas with lower median household income do not necessarily have greater access to the program and that spatial disparities still persist. Under the the conditions of the original system, which did not have equity stations, _X%_ of dissemination areas (or _X%_ of the population) within the core service area with median household income less than 40,000 Canadian dollars could access at least one bicycle share station within 10 minutes walk. Only _X%_ of dissemination areas (or _X%_ of the population) with median household income less than 40,000 Canadian dollars could access at least one bicycle share station within 3 or 5 minutes walk.

Similar to other studies, we find that areas with higher median household income have greater accessibility to Hamilton Bike Share. Our findings align with other studies that have found disparities in station or bike location along income lines in New York City [@babagoliExploringHealthSpatial2019], Tampa [@chenExploringEquityPerformance2019], and Seattle [@mooneyFreedomStationSpatial2019].

# Study Limitations

>>>>

```{r sobi-stations-in-hamilton, echo=FALSE, fig.align = 'center', out.width = "100%", fig.cap = "The spatial distribution of bike share stations in Hamilton, Ontario. The service area of Hamilton Bike Share is outlined in blue and the city's downtown core is outlined in light blue. Dissemination areas within the Hamilton CMA that touch the bounding box of the core service area are outlined in grey."}
ggplot() + 
  geom_sf(data = hamilton_da_2016 %>% 
            dplyr::filter(sobi_service == "Yes"),
          color = "gray") +
  geom_sf(data = sobi_service, 
          fill = NA,
          color = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = "dark green",
          color = "dark green") +
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional" & 
                            hub_status == "Active"), # Remove deactivated hubs
          aes(size = RACKS_AMOU),
          shape = 21,
          fill = "blue") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          aes(size = RACKS_AMOU), 
          shape = 21,
          fill = "green")
```

@@

## Accessibility by Distance

Consensus regarding the distance that individuals are willing to walk to access a PBSP is lacking, therefore the literature on walking behaviour was consulted to determine the thresholds for the sensitivity analysis. Previous studies have found that living within 250 metres [@fullerUseNewPublic2011] and 300 metres [@kabraBikeShareSystemsAccessibility2020] is correlated with bike share use, while other research has found that walking trips are less than 600 metres and rarely more than 1200 metres [@millwardActivetransportWalkingBehavior2013] or a median distance of 650 metres [@larsenQuarterMileReexamining2010]. As seen in Table \ref{tab:accessibility-sensibility}, we find that accessibility increases with a threshold between two and four minutes, but is then maximized at 5 minutes. At some stations, Hamilton Bike Share will depict a map to show the user the locations of the other nearest stations within a five minute walk, which suggests that this is an average distance that people are anticipated to walk. Accessibility decreases significantly after eight minutes, which is intuitive given that demand on a limited supply increases as more people can reach each bike share station.

For this reason, we experiment with various walking thresholds by conducting a sensitivity analysis to calculate accessibility at different walking times: 3 minutes, 5 minutes, 10 minutes, and 15 minutes. We categorize these thresholds as minimum, average, maximum, and extreme, respectively. At each threshold, we compare accessibility between the current system and the original system to examine the contribution of the additional equity stations.

```{r accessibility-sensibility, echo=FALSE}
# Testing different thresholds to check how accessibility changes
accessibility_test <- data.frame(threshold = c(3:15), 
                                 accessibility = numeric(13))
for(i in 1:13){
  results_test <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                hub_status == "Active") %>%
  b2sfca(threshold = i +2)
  accessibility_test$accessibility[i] <- sum(results_test$accessibility$accessibility)
}

ggplot(accessibility_test, 
       aes(x = threshold, y = accessibility)) +
  geom_line()
```

```{r join-population-to-travel-time-data, include=FALSE}
# current system with equity stations
ttm_walk <- left_join(ttm_walk, 
                      population_50x50 %>%
                        st_drop_geometry(),
                      by = "UID")
```

@@

```{r, include=FALSE}
#data frames for current and original systems
names(ttm_walk) <- c("UID", 
                     "hub", 
                     "travel_time",
                     "hub_type",
                     "hub_status",
                     "population",
                     "racks")
```

### Three Minutes

At this threshold, we find that there are 25.2 bicycle racks per person in the original system.  With the addition of equity stations, there are now 25.4 bicycles per person. Figure \ref{fig:figure-6} presents a comparison of accessibility between the systems. Accessibility is fairly uniform, with the exception of two small areas near the university and the waterfront park where accessibility is slightly higher. Equity stations are shown in green, which helps to demonstrate how they increase accessibility between the systems.

```{r min-conventional, include=FALSE}
results_3min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

```{r min-all, include=FALSE}
results_3min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

```{r min-sum-conventional, echo=FALSE}
sum(results_3min_conventional$los$los)
sum(results_3min_conventional$accessibility$accessibility)
```

```{r min-sum-all, echo=FALSE}
sum(results_3min_all$los$los)
sum(results_3min_all$accessibility$accessibility)
```
```{r figure-6, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 3 minutes walk (minimum threshold) compared between current system with equity stations and the original system without equity stations."}
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "3 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "PuRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

### Five Minutes

At this threshold, we find that there are 68.6 bicycle racks per person in the original system.  With the addition of equity stations, there are now 68.8 bicycles per person. Figure \ref{fig:figure-7} presents a comparison of accessibility between the systems. Again, accessibility is fairly uniform, with the exception of one area near the waterfront park where accessibility is higher.

```{r med-conventional, include=FALSE}
results_5min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

```{r med-all, include=FALSE}
results_5min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

```{r, med-sum-conventional, echo=FALSE}
sum(results_5min_conventional$los$los)
sum(results_5min_conventional$accessibility$accessibility)
```

```{r med-sum-all, echo=FALSE}
sum(results_5min_all$los$los)
sum(results_5min_all$accessibility$accessibility)
```

```{r figure-7, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 5 minutes walk (average threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 5 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "5 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "PuRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

### Ten Minutes

At this threshold, we find that there are 3.61 bicycle racks per person in the original system.  With the addition of equity stations, there are now 3.74 bicycles per person. Figure \ref{fig:figure-8} presents a comparison of accessibility between the systems. We begin to see differences in accessibility across the service area, with users near the university and its adjacent neighbourhoods, as well as neighbourhoods north of the downtown area have slightly higher accessibility. While the differences are modest, they are more apparent at this threshold than at shorter walking distances.

```{r avg-conventional, include=FALSE}
results_10min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

```{r avg-all, include=FALSE}
results_10min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

```{r avg-sum-conventional, echo=FALSE}
sum(results_10min_conventional$los$los)
sum(results_10min_conventional$accessibility$accessibility)
```

```{r, avg-sum-all, echo=FALSE}
sum(results_10min_all$los$los)
sum(results_10min_all$accessibility$accessibility)
```

```{r figure-8, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 10 minutes walk (maximum threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 10 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "10 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "PuRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

### Fifteen Minutes

At this threshold, we find that there are 2.44 bicycle racks per person in the original system.  With the addition of equity stations, there are now 2.55 bicycles per person. Figure \ref{fig:figure-9} presents a comparison of accessibility between the systems. Users near the university and the neighbourhoods north of the downtown area (outlined in green) have the highest accessibility, followed by those who live in the city's downtown area. Accessibility in the east end of the core service area remains low.

```{r max-conventional, include=FALSE}
results_15min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

```{r max-all, include=FALSE}
results_15min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

```{r max-sum-conventional, echo=FALSE}
sum(results_15min_conventional$los$los)
sum(results_15min_conventional$accessibility$accessibility)
```

```{r max-sum-all, echo=FALSE}
sum(results_15min_all$los$los)
sum(results_15min_all$accessibility$accessibility)
```

```{r figure-9, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 15 minutes walk (extreme threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 15 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "15 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "PuRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

```{r, include=FALSE}
los <- rbind(data.frame(results_3min_all$los, system = "All stations", threshold = "3 min"),
             data.frame(results_3min_conventional$los, system = "Conventional", threshold = "3 min"),
             data.frame(results_5min_all$los, system = "All stations", threshold = "5 min"),
             data.frame(results_5min_conventional$los, system = "Conventional", threshold = "5 min"),
             data.frame(results_10min_all$los, system = "All stations", threshold = "10 min"),
             data.frame(results_10min_conventional$los, system = "Conventional", threshold = "10 min"),
             data.frame(results_15min_all$los, system = "All stations", threshold = "15 min"),
             data.frame(results_15min_conventional$los, system = "Conventional", threshold = "15 min"))

@@

## Accessibility by Median Total Income

Approximately 138,000 people live within Hamilton Bike Share's core service area or within a 10 minute walk to the service area. We find that over 118,000 people can access a bicycle share station within a 15 minute walk, which represents roughly 85% of the total population in Hamilton Bike Share's core service area [see Table \ref{tab:table-2}]. While previous research found that neighbourhoods with more disadvantage are better serviced by Hamilton Bike Share [@hosfordWhoArePublic2018], the authors used the Pampalon Deprivation Index from 2011 as a measure of socioeconomic status. Instead, we use median total income for each dissemination area to examine whether disparities in accessibility exist within the core service area. We divide income by quantiles: bottom 20%, second 20%, third 20%, fourth 20%, and top 20%.  One of the unique properties of the balanced floating catchment area method is that data can be reaggregated from small to larger polygons while preserving the total population and number of bicycle racks at each station. This avoids demand and supply inflation, and also enables us to present findings in a way that is easier to interpret (i.e., at the dissemination area level).

Panels 1 and 2 depict bivariate choropleth maps with the combined spatial distribution of accessibility and median total income per household for the different thresholds. Our analysis demonstrates that the stations added to Hamilton's public bicycle share program to increase equity for disadvantaged neighbourhoods achieved their goal by increasing accessibility, albeit only modestly. By implementing equity stations in more areas with lower median total income in Hamilton, the PBSP has achieved greater horizontal equity by extending the spatial distribution of bicycles across the city. This is particularly evident at shorter walking distances of three and five minutes.

However, we found that dissemination areas with lower total household income do not necessarily have greater access to the program [see Table \ref{tab:table-3}]. Income disparities still persist, however only at certain thresholds. With and without equity stations, people in the top 20% of income have the highest level of accessibility at a threshold of ten and fifteen minutes. At distance thresholds of three and five minutes, people in the second 20% have the highest level of accessibility by a significant amount. This is due to the concentration of stations near or within the city's downtown area and the "lower city" which historically has been more disadvantaged compared to the surrounding suburban neighbourhoods. This is an artifact of the city's steel industry which led to more blue collar and working class people living in the city's downtown area and "lower city". However, the bottom 20%, who may benefit the most from Hamilton Bike Share's equity initiatives, have the lowest accessibility at three minutes threshold and the second lowest accessibility at all other thresholds.

On the whole, we find that areas that have less than their equitable share of the level of service are some of the most socioeconomically disadvantaged areas in the city. This finding aligns with other studies that have found disparities in station or bike location along income lines in New York City [@babagoliExploringHealthSpatial2019], Tampa [@chenExploringEquityPerformance2019], and Seattle [@mooneyFreedomStationSpatial2019].

```{r, include=FALSE}
# drop geometry
test <- accessibility

# join accessibility to population cells
test <- test %>% st_join(population_50x50 %>% dplyr::select(-UID)) %>%
  rename(population50x50 = population)
```

```{r, include=FALSE}
test <- test %>%
  st_join(hamilton_da_2016)
```

```{r, include=FALSE}
agg_accessibility_da <- test %>%
  group_by(DA, 
           system,
           threshold) %>%
  summarize(accessibility = sum(accessibility),
            median_total_income = first(median_total_income), 
            .groups = "drop") %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% dplyr::select(DA, geometry),
            by = c("DA" = "DA")) %>%
  st_as_sf()
```

```{r figure-10, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "3 min", 
                          system == "All stations") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkCyan", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, 
          shape = 21, 
          fill = "orange") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income with Equity Stations \n at 3 Minutes Threshold") + 
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkCyan",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(data_map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-11, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "3 min", 
                          system == "Conventional") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkCyan", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income without Equity Stations \n at 3 Minutes Threshold") + 
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkCyan",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(data_map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-12, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "5 min", 
                          system == "All stations") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkCyan", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, 
          shape = 21, 
          fill = "orange") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income with Equity Stations \n at 5 Minutes Threshold") + 
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkCyan",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(data_map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-13, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "5 min", 
                          system == "Conventional") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkCyan", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income without Equity Stations \n at 3 Minutes Threshold") + 
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkCyan",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(data_map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-14, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "10 min", 
                          system == "All stations") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkCyan", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, 
          shape = 21, 
          fill = "orange") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income with Equity Stations \n at 10 Minutes Threshold") + 
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkCyan",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(data_map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-15, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "10 min", 
                          system == "Conventional") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkCyan", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income without Equity Stations \n at 10 Minutes Threshold") +
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkCyan",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(data_map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-16, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "15 min", 
                          system == "All stations") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, 
          shape = 21, 
          fill = "orange") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income with Equity Stations \n at 15 Minutes Threshold") + 
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(data_map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-17, echo=FALSE, fig.align = 'center', out.width = "65%"}
data_map <- agg_accessibility_da %>%
            dplyr::filter(threshold == "15 min", 
                          system == "Conventional") %>%
  bi_class(x = accessibility, 
           y = median_total_income, 
           style = "quantile", 
           dim = 3) %>%
  st_as_sf()

data_map <- ggplot() +
  geom_sf(data = data_map,
          mapping = aes(fill = bi_class), 
          color = "white", 
          size = 0.1, 
          show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", 
                dim = 3) +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "purple") +
  bi_theme() +
  ggtitle("Accessibility by Income without Equity Stations \n at 15 Minutes Threshold") + 
  theme(plot.title = element_text(face = "bold", colour = "black", size = 12))

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)
